---
title: "recruitment_brightspots"
author: "Christopher Rounds"
date: "11/17/2023"
output: html_document
---

```{r}
library(tidyverse)
library(mnsentinellakes)
library(arrow)
library(sf)
library(mgcv)

source("rf_functions.r")

mn_lakes <- readRDS("./data/covariate/mndow_lakes_sf_allDataUntransformed.rds")

mn_data <- open_dataset("./data/mn_file_arrow/") 

efish_wae <- read.csv("./data/Kundel_EF_DATA_Final.csv") %>% 
  mutate(DOW = fixlakeid(DOW_KITTLE)) %>%
  dplyr::select(!c(SURVEY_ID, DOW_KITTLE, AREA_NAME, REGION_NAME, SAMPLING_END_DATE)) %>%
  dplyr::select(c(DOW, WATER_BODY_NAME, SURVEY_YEAR, TOTAL_CATCH), everything())

  
mw_data <- open_dataset("./data/hive", partitioning = c("state"))
```


# Collect Fall electrofishing for walleye surveys
```{r}
surveys <- mw_data %>% filter(state == "Minnesota") %>% 
  filter(sampling_method == "Fall electrofishing for walleye") %>%
  mutate(lubridate_date_clean = lubridate::as_date(date_sample),
         year = year(lubridate_date_clean), month = lubridate::month(lubridate_date_clean)) %>%
  dplyr::filter(total_effort_1 > 0.5) %>%
  collect()

non_zero <- surveys %>%
  dplyr::filter(species_1 == "walleye") %>%
  dplyr::filter(age_class == "young_of_year") %>%
  group_by(lake_id, lake_name, nhdhr_id, total_effort_ident, water_temp,
           date_survey, date_total_effort_ident, date_sub_effort_ident, date_sample, 
           year, total_effort_1) %>%
  summarize(catch = n()) %>% mutate(CPUE = catch/total_effort_1)

zero_surveys = surveys %>%
  dplyr::filter(!total_effort_ident %in% non_zero$total_effort_ident) %>%
  group_by(lake_id, lake_name, nhdhr_id, total_effort_ident, water_temp,
           date_survey, date_total_effort_ident, date_sub_effort_ident, date_sample, 
           year, total_effort_1) %>%
  summarize(catch = 0) %>% mutate(CPUE = 0) 

new_surveys_temp <- rbind(non_zero, zero_surveys) %>% 
  ungroup() %>%
  mutate(julian_day_survey = yday(date_survey),
         julian_day_sample = yday(date_total_effort_ident),
         julian_day_survey = ifelse(julian_day_sample > julian_day_survey, 
                                    julian_day_sample, julian_day_survey)) %>%
  mutate(lake_id = fixlakeid(lake_id),
         method_abbrev = "EW")

notes <-surveys %>% filter(total_effort_ident %in% new_surveys_temp$total_effort_ident) %>% 
  select(total_effort_ident, gear_data_notes) %>% group_by(total_effort_ident) %>%
  slice(1) %>% ungroup()

new_surveys_ew <- merge(new_surveys_temp, notes, by = "total_effort_ident")
```

# Collect Standard electrofishing surveys
```{r}
surveys <- mw_data %>% filter(state == "Minnesota") %>% 
  filter(sampling_method == "Standard electrofishing" | 
           sampling_method == "Special sampling, electrofishing") %>%
  filter(target_species == "walleye") %>%
  mutate(lubridate_date_clean = lubridate::as_date(date_sample),
         year = year(lubridate_date_clean), month = lubridate::month(lubridate_date_clean)) %>%
  dplyr::filter(total_effort_1 > 0.5) %>%
  collect()

non_zero <- surveys %>%
  dplyr::filter(species_1 == "walleye") %>%
  dplyr::filter(age_class == "young_of_year") %>%
  group_by(lake_id, lake_name, nhdhr_id, total_effort_ident, water_temp,
           date_survey, date_total_effort_ident, date_sub_effort_ident, date_sample, 
           year, total_effort_1) %>%
  summarize(catch = n()) %>% mutate(CPUE = catch/total_effort_1)

zero_surveys = surveys %>%
  dplyr::filter(!total_effort_ident %in% non_zero$total_effort_ident) %>%
  group_by(lake_id, lake_name, nhdhr_id, total_effort_ident, water_temp,
           date_survey, date_total_effort_ident, date_sub_effort_ident, date_sample, 
           year, total_effort_1) %>%
  summarize(catch = 0) %>% mutate(CPUE = 0) 

new_surveys_temp <- rbind(non_zero, zero_surveys) %>% 
  ungroup() %>%
  mutate(julian_day_survey = yday(date_survey),
         julian_day_sample = yday(date_total_effort_ident),
         julian_day_survey = ifelse(julian_day_sample > julian_day_survey, 
                                    julian_day_sample, julian_day_survey)) %>%
  mutate(lake_id = fixlakeid(lake_id),
         method_abbrev = "EF")

notes <-surveys %>% filter(total_effort_ident %in% new_surveys_temp$total_effort_ident) %>% 
  select(total_effort_ident, gear_data_notes) %>% group_by(total_effort_ident) %>%
  slice(1) %>% ungroup()

new_surveys_sef_ef <- merge(new_surveys_temp, notes, by = "total_effort_ident")
```

# Combine all three
```{r}
new_surveys <- rbind(new_surveys_ew, new_surveys_sef_ef) %>%
  filter(julian_day_survey > 213) %>%
  filter(!str_detect(gear_data_notes, "DAYLIGHT_SAMPLING:Y")) %>%
  mutate(year = year(date_survey))

new_surveys %>% group_by(lake_id, year) %>% count() %>% filter(n >1)

# remove bad duplicate surveys within a lake-year
bad_efforts = c("335", "6887", "6881", "6877", "6874", "11867", "18106", "18341", 
                "20442", "20444", "20445", "20446", "20447", "21402", "21404" , 
                "21407", "21416", "21465", "21449" ,"22077", "23736", "31943", 
                "32013", "32009", "35885", "40686", "43225")

new_surveys <- new_surveys %>% filter(!total_effort_ident %in% bad_efforts)
new_ef_dows <- unique(new_surveys$lake_id)
```


```{r}
rm(new_surveys_ew); rm(surveys); rm(new_surveys_sef_ef); rm(notes); rm(new_surveys_temp);
rm(non_zero); rm(zero_surveys)
```


# Get lake attributes for the new surveys
```{r}
sf_use_s2(FALSE)
st_crs(mn_lakes) <- 26915

mn_lakes_ef <- mn_lakes %>% dplyr::filter(dowlknum %in% new_ef_dows)

mn_lakes_ef <- st_as_sf(x = mn_lakes_ef,                         
                  coords = c("x", "y"),
                  crs = "+proj=utm +zone=15")
mn_lakes_ef = st_transform(mn_lakes_ef, crs = "+proj=longlat +datum=WGS84")
mn_lakes_ef$x <- st_coordinates(st_centroid(st_as_sf(mn_lakes_ef)))[,"X"]
mn_lakes_ef$y <- st_coordinates(st_centroid(st_as_sf(mn_lakes_ef)))[,"Y"]

mn_lakes <- as.data.frame(mn_lakes_ef) %>% 
    # make sure some riverish lakes stay (lake pepin, laq que parle)
  mutate(wb_class = ifelse(dowlknum == "25000100", 
                           "semiriverine polygon", wb_class),
         wb_class = ifelse(dowlknum == "37004600", 
                           "semiriverine polygon", wb_class),
         wb_class = ifelse(dowlknum == "62004700", 
                           "semiriverine polygon", wb_class),
         wb_class = ifelse(dowlknum == "69129100", 
                           "semiriverine polygon", wb_class),
         wb_class = ifelse(dowlknum == "69129100", 
                           "semiriverine polygon", wb_class),
         wb_class = ifelse(dowlknum == "79000100", 
                           "semiriverine polygon", wb_class)) %>%
  select("dowlknum", "pw_basin_n","acres", "shore_mi", "x", "y", "wb_class") %>%
  rename(lake_id = dowlknum) %>% distinct(lake_id, .keep_all = T)

new_surveys_spatial <- merge(x = new_surveys, y = mn_lakes, by = "lake_id", all.x = T) %>%
  # I dont think is needed anymore
  dplyr::filter(wb_class != "Riverine polygon")
```


# Get invasion status for all lakes
```{r}
ais <- readxl::read_excel("./data/covariate/infested-waters.xlsx") %>% 
  dplyr::filter(ais == "zebra mussel") %>% select(!Column1) %>%
  dplyr::filter(dow != "NA", dow != "na") %>%
  dplyr::filter(!grepl("none", dow) ) %>%
  mutate(lake_id = fixlakeid(dow),
         lake_id = str_pad(lake_id, width = 8, side = "right", pad = "0")) %>%
  dplyr::rename(year.listed = year) %>%
  select(lake_id, year.listed, year_confirmed, ais)

new_surveys_ais <- merge(x = new_surveys_spatial, y = ais, 
                         by = "lake_id", all.x = T) %>%
  mutate(ais = ifelse(is.na(ais),0, 1),
         year_confirmed = ifelse(is.na(year_confirmed), 0, year_confirmed),
         infested = as.factor(ifelse(year_confirmed >= year, 1, 0)))


new_surveys_ais <- merge(x = new_surveys_spatial, y = ais, 
                         by = "lake_id", all.x = T) %>%
  mutate(ais = ifelse(is.na(ais), 0, 1),
         year_confirmed = as.numeric(year_confirmed),
         year_confirmed = ifelse(is.na(year_confirmed), NA, year_confirmed),
         infested = ifelse(year_confirmed <= year, 1, 0),
         BACI = ifelse(year_confirmed >= year, 1, 0),
         BACI = ifelse(infested == 1, 2, BACI),
         BACI = ifelse(is.na(BACI), 3, BACI),
         BACI = ifelse(BACI == 3 & year > 2016, 4, BACI)) %>%
  dplyr::select(!c(year.listed, ais))
#BACI - 1 = infested lake before infestation
#BACI - 2 = Infested lake post infestation
#BACI - 3 = uninfested lake pre-pseudo year
#BACI - 4 = uninfested lake post pseudo year
```


# Get new stocking 
```{r}
stocking <- read.csv("./data/covariate/mn_wae_stocking.csv") %>% 
  dplyr::filter(Year > 2019) %>% rename(lake_id = DOW, year = Year) %>%
  mutate(lake_id = ifelse(lake_id == "04003500", "04003501", lake_id))

new_surveys_stocking <- merge(x = new_surveys_ais, y = stocking, 
                         by = c("lake_id", "year"), all.x = T) %>%
  mutate(across(FRY:frl.pa, ~ ifelse(is.na(.), 0, .))) %>%
  #If any are stocked make Stocked = 1
  mutate(STOCKED =  ifelse(rowSums(across(fry.pa:frl.pa)) > 0, 1, 0))
```


# Get Secchi
```{r}

secchi <- read.csv("./data/covariate/MN_AnnualSecchi_update.csv") %>%
  select(year, median_secchi, Prmnn_I) %>%
  mutate(nhdhr.id = paste0("nhdhr_", Prmnn_I))


nhd.id <- mwlaxeref::lake_id_xref %>% filter(state == "mn") %>%
  mutate(local.id = fixlakeid(local.id)) %>%
  group_by(local.id, nhdhr.id) %>% distinct(.keep_all = T) %>% ungroup()

secchi.full = merge(x = secchi, y = nhd.id, 
               by = c("nhdhr.id"), all.x = T) %>%
  select(local.id, nhdhr.id, lagos.id, year, median_secchi) %>%
  rename(lake_id = local.id) %>%
  group_by(lake_id, nhdhr.id, year) %>% 
  distinct(.keep_all = T) %>% ungroup()

# add remotely sensed secchi to the gill net data
new_surveys_secchi <- merge(x = new_surveys_stocking, y = secchi.full, 
                   by = c("lake_id", "year"), all.x = T) 

new_surveys_secchi <- new_surveys_secchi %>% group_by(year, lake_id) %>%
  mutate(median_secchi = mean(median_secchi)) %>% 
  distinct(total_effort_ident, .keep_all = T)
```


# Merge with big dataframe
```{r}
new_surveys_merge <- new_surveys_secchi %>%
    select(year, lake_id, catch, total_effort_1, acres, shore_mi, x, y, infested, 
           year_confirmed, julian_day_survey, STOCKED, median_secchi) %>% #select(!adl.pa) %>%
  mutate(STOCKED = as.factor(STOCKED)) %>% ungroup()

surveys.ef <- new_surveys_merge %>%
  mutate(year_confirmed = as.numeric(year_confirmed),
         year_confirmed = ifelse(is.na(year_confirmed),0, year_confirmed),
         BACI = ifelse(year_confirmed > year, 1, 0),
         BACI = ifelse(year_confirmed > 1900 & year_confirmed <= year, 2, BACI),
         BACI = ifelse(year_confirmed < 1900, 3, BACI),
         BACI = ifelse(BACI == 3 & year > 2016, 4, BACI)) 

stocking <- read.csv("./data/covariate/mn_wae_stocking.csv") %>% 
  rename(lake_id = DOW, year = Year) %>%
  # Fix red lake stocking
  mutate(lake_id = ifelse(lake_id == "04003500", "04003501", lake_id))


surveys_stocking <- merge(x = surveys.ef, y = stocking, 
                         by = c("lake_id", "year"), all.x = T) %>%
  mutate(across(FRY:frl.pa, ~ ifelse(is.na(.), 0, .))) %>%
  #If any are stocked make Stocked = 1
  mutate(STOCKED =  ifelse(rowSums(across(fry.pa:frl.pa)) > 0, 1, 0),
         fry.pa = FRY/acres, 
         fgl.pa = FGL/acres, 
         adl.pa = ADL/acres, 
         yrl.pa = YRL/acres,
         frl.pa = FRL/acres) %>%
  dplyr::select(-c(FRY:FRL))
```


# Remove old objects
```{r}
rm(new_surveys); rm(new_surveys_ais); rm(new_surveys_spatial); rm(new_surveys_stocking)
rm(ais); rm(stocking); rm(mn_lakes);rm(efish_wae);
rm(secchi.full); rm(secchi); rm(surveys.ef)
rm(new_surveys_merge); rm(new_surveys_secchi); rm(nhd.id)
```


# Get Temp data
```{r}
current_temps <- read_feather("./data/lake_temperature_metrics_GLM_NLDAS.feather") %>% 
  dplyr::select(site_id, year, peak_temp, ice_off_date, coef_var_1_30, coef_var_31_60,
                gdd_wtr_0c, gdd_wtr_5c, gdd_wtr_10c, contains("mean_surf"),
                post_ice_warm_rate, contains("date_over"))


dow_year <- surveys_stocking %>% distinct(lake_id, year)
dow_nhdhr <- readRDS("./data/mndow_nhdhr_xwalk.rds") %>% rename(lake_id = MNDOW_ID) %>%
  mutate(lake_id = gsub("^.{0,6}", "", lake_id))

missing_ids <- data.frame(
  lake_id = c("69069400", "69060800", "38053200", "38021100", "16063300"), 
  site_id = c("nhdhr_120019354", "nhdhr_80993113", 
               "nhdhr_72d19d48-754d-459b-a1f2-b1cfd8993b06", "nhdhr_80993749", 
               "nhdhr_120019082"))
dow_nhdhr = dow_nhdhr %>% rows_update(y = missing_ids, by = c("lake_id"))

surveys_no_temp <- merge(surveys_stocking, dow_nhdhr, by = c("lake_id"), all.x = T)

filtered_temp <- current_temps %>% 
  dplyr::filter(site_id %in% surveys_no_temp$site_id) 

surveys_ef <- merge(x = surveys_no_temp , y = filtered_temp, 
                    by = c("site_id", "year"), all.x = T)
```

# remove temperature stuff
```{r}
rm(current_temps); rm(filtered_temp); rm(surveys_no_temp); rm(dow_nhdhr); rm(dow_year); 
rm(mn_lakes_ef)
```

# Final Fixes before CSVing
```{r}
model_surveys_ef <- surveys_ef %>% 
  mutate(yrl.pa = ifelse(is.na(yrl.pa), 0, yrl.pa),
         log_acres = log(acres),
         log_shore = log(shore_mi),
         cpue = catch/total_effort_1,
         ice_off_yday = yday(ice_off_date)) %>%
  dplyr::select(-ice_off_date) %>%
  mutate(infested = ifelse(is.na(infested), 0, infested)) %>% drop_na()
  group_by(lake_id) %>%
  mutate(median_secchi = ifelse(is.na(median_secchi), 
                                median(median_secchi,na.rm = T), 
                                median_secchi)) %>%
  mutate(across(where(is.numeric), ~ 
                  replace_na(., median(., na.rm = TRUE)))) %>%
  ungroup() %>% drop_na()


yoy_ef <- model_surveys_ef %>%
  select(!c(site_id, acres, shore_mi,
            infested, fgl.pa, adl.pa, yrl.pa, frl.pa))

#write.csv(yoy_ef, "./data/yoy_catch_data.csv", row.names = F)
```




### GRAVEYARD

# Get new EF surveys
```{r}
# collect all relevant surveys
recent_surveys = mn_data %>%
  filter(sampling_method_abbrev %in% c("EW", "EF", "SEF")) %>%
  mutate(lubridate_date_clean = lubridate::as_date(date_clean),
         year = year(lubridate_date_clean)) %>%
  dplyr::filter(year > 2019) %>%
  dplyr::filter(month(lubridate_date_clean) >= 8) %>%
  dplyr::filter(total_effort_1.1 > 0.5) %>%
  dplyr::filter(!str_detect(target_species, "bass|carp|perch|bluegill|muskellunge")) %>%
  distinct(lake_id, nhdhr.id,lake_name.1, date_clean, year, 
           total_effort_1.1, effort_ident) %>%
  #dplyr::filter(young_of_year == "Y") %>%
  collect()


# collect all relevant fish information
recent_efishing = mn_data %>%
  filter(sampling_method_abbrev %in% c("EW", "EF", "SEF"), 
         species_abbrev %in% c("WAE")) %>%
  mutate(lubridate_date_clean = lubridate::as_date(date_clean),
         year = year(lubridate_date_clean)) %>%
  dplyr::filter(year > 2019) %>%
  dplyr::filter(month(lubridate_date_clean) >= 8) %>%
  dplyr::filter(total_effort_1.1 > 0.5) %>%
  dplyr::filter(young_of_year == "Y") %>%
  dplyr::filter(!str_detect(target_species, "bass|carp|perch")) %>%
  collect()


new_efishing <- recent_efishing %>% 
  group_by(lake_id, nhdhr.id, lake_name.1, effort_ident, 
           date_clean, year, total_effort_1.1) %>%
  summarize(catch = n()) %>% mutate(CPUE = catch/total_effort_1.1)

# Find surveys where no WAE YOY caught
zero_catch <- recent_surveys %>% 
  dplyr::filter(!effort_ident %in% new_efishing$effort_ident) %>%
  mutate(catch = 0, CPUE = 0)

new_surveys <- rbind(new_efishing, zero_catch) %>%
  mutate(lake_id = fixlakeid(lake_id)) %>%
  #junky surveys
  filter(!effort_ident %in% c("4600", "336", "28882", "14566"))

new_ef_dows <- unique(new_surveys$lake_id)

rm(zero_catch); rm(new_efishing); rm(recent_efishing)
```


# Attempting to work on recruitment problems
```{r, eval = F}


temp_e = mn_data %>% 
  filter(sampling_method_abbrev %in% c("EW", "EF", "SEF")) %>%
  mutate(lubridate_date_clean = lubridate::as_date(date_clean),
         year = year(lubridate_date_clean), month = lubridate::month(lubridate_date_clean)) %>%
  dplyr::filter(!str_detect(target_species, 
                            "bass|carp|perch|bluegill|muskellunge|catfish|crappie|sturgeon")) %>%
  dplyr::filter(total_effort_1.1 > 0.5) %>%
  collect() %>%
  dplyr::filter(species.1 == "walleye") %>%
  dplyr::filter(young_of_year == "Y") %>%
  group_by(lake_id, lake_name.1, nhdhr.id, effort_ident, date_clean, year, total_effort_1.1) %>%
  summarize(catch = n()) %>% mutate(CPUE = catch/total_effort_1.1)


surveys_certain_missing = mn_data %>% 
  filter(sampling_method_abbrev %in% c("EW")) %>%
  mutate(lubridate_date_clean = lubridate::as_date(date_clean),
         year = year(lubridate_date_clean), month = lubridate::month(lubridate_date_clean)) %>%
  dplyr::filter(lake_id %in% c(69037800, 25000100)) %>%
  dplyr::filter(!str_detect(target_species, 
                            "bass|carp|perch|bluegill|muskellunge|catfish|crappie|sturgeon")) %>%
  dplyr::filter(total_effort_1.1 > 0.5) %>%
  collect() %>%
  dplyr::filter(species.1 == "walleye") %>%
  dplyr::filter(young_of_year == "Y") #%>%
  mutate(uid = paste0(lake_id, year, total_effort_1.1, sampling_method_abbrev)) %>%
  group_by(lake_id, lake_name.1, nhdhr.id, effort_ident, date_clean, year, total_effort_1.1, uid) %>%
  summarize(catch = n()) %>% mutate(CPUE = catch/total_effort_1.1)

# U think this is wrong 
## Lots of walleyes caught but all marked as NON-YOY even though many are not lengthed
surveys_certain_missing_zero = mn_data %>% 
  filter(sampling_method_abbrev %in% c("EW")) %>%
  mutate(lubridate_date_clean = lubridate::as_date(date_clean),
         year = year(lubridate_date_clean), month = lubridate::month(lubridate_date_clean)) %>%
  dplyr::filter(lake_id %in% c(69037800, 25000100)) %>%
  dplyr::filter(!str_detect(target_species, 
                            "bass|carp|perch|bluegill|muskellunge|catfish|crappie|sturgeon")) %>%
  dplyr::filter(total_effort_1.1 > 0.5) %>%
  collect() %>%
  dplyr::filter(!effort_ident %in% surveys_certain_missing$effort_ident) %>%
  mutate(uid = paste0(lake_id, year, total_effort_1.1, sampling_method_abbrev)) %>%
  dplyr::filter(!uid %in% surveys_certain_missing$uid) %>%
  group_by(lake_id, lake_name.1, nhdhr.id, effort_ident, date_clean, year, total_effort_1.1) %>%
  summarize(catch = n()) %>%
  mutate(catch = 0, CPUE = 0)



new_surveys <- surveys_certain_missing %>% select(!uid) %>% 
  #rbind(surveys_certain_missing_zero) %>% 
  mutate(lake_id = as.character(lake_id)) %>%
  rbind(new_surveys)
new_ef_dows <- unique(new_surveys$lake_id)



efish_merge <- efish_wae %>% 
  mutate(julian_day = yday(as.Date(SAMPLING_START_DATE)),
         acres = AREA_HECTARES*2.47105,
         YEAR_INFESTED = ifelse(is.na(YEAR_INFESTED), 0, YEAR_INFESTED),
         infested = as.factor(INVASION_STATUS_BINARY)) %>%
  select(SURVEY_YEAR, DOW, TOTAL_CATCH, EFFORT, acres, SHORE_LENGTH_MILES,
         STOCKED, julian_day, YEAR_INFESTED,
         infested, LAKE_CENTER_LAT_DD5, LAKE_CENTER_LONG_DD5, 
         ANNUAL_MED_SECCHI_M) %>%
  rename(year = SURVEY_YEAR, lake_id = DOW, 
         total_effort_1.1 = EFFORT, catch = TOTAL_CATCH,
         shore_mi = SHORE_LENGTH_MILES, median_secchi = ANNUAL_MED_SECCHI_M,
        year_confirmed = YEAR_INFESTED, 
         x = LAKE_CENTER_LONG_DD5, y = LAKE_CENTER_LAT_DD5) 
```

