---
title: "rf_adult"
output: html_document
---

```{r setup, include=FALSE}
library(rfUtilities)
library(pdp)
library(randomForest)
library(tidyverse)
library(lme4)
library(partykit)
source("rf_functions.r")
options(scipen=999)

# gill net data
gn_temp_train <- read.csv("./data/adult_catch_data.csv") %>%
  mutate(BACI = as.factor(BACI),
         STOCKED = as.factor(STOCKED),
         ice_off_date = lubridate::yday(ice_off_date),
         ice_off_date.previous = lubridate::yday(ice_off_date.previous),
         ice_off_date.2.previous = lubridate::yday(ice_off_date.2.previous),
         log_cpue = log(cpue + 0.01), 
         three_year_secchi = (median_secchi + median_secchi_previous +
                                median_secchi_previous_2)/3)

gn_temp_train <- gn_temp_train[complete.cases(gn_temp_train), ]

# E-fishing YOY data
yoy_surveys <- read.csv("./data/yoy_catch_data.csv") %>%
    mutate(BACI = as.factor(BACI),
         STOCKED = as.factor(STOCKED),
         log_cpue = log(cpue + 0.01))

# size structure data
size_temp_train <- read.csv("./data/adult_size_data.csv") %>% 
  mutate(BACI = as.factor(BACI),
         STOCKED = as.factor(STOCKED),
         three_year_secchi = (median_secchi + median_secchi_previous +
                                median_secchi_previous_2)/3)
size_temp_train <- size_temp_train[complete.cases(size_temp_train), ]
```


# Gill net modeling
```{r}
independent.gn <- gn_temp_train %>% 
  select(!c("lake_id", "lake_name.1", "cpue", 
            "count", "total_effort_1.1", "x", "y"))


cor <- Hmisc::rcorr(as.matrix(independent.gn))$r

# high correlation (>.6)
# acres-shore miles, secchi has ~.4-.5 with all temp metrics
# practically all temp metrics
# good temp variables post-ice-warm rate, coef_var, date_over

independent.gn <- gn_temp_train %>% 
  select(c("year", "julian_day", "cpue_yep","cpue_np","cpue_cisco",
           "log_acres", "max_depth", "elevation", "BACI", 
           "three_year_secchi",
           "STOCKED", "STOCKED.previous", "STOCKED.2.previous", 
           "fry.pa", "fry.pa.previous", "fry.pa.2.previous",
           "gdd_wtr_0c", "post_ice_warm_rate", "date_over_16.7", 
           "mean_surf_JulAugSep","mean_surf_may", "winter_dur_0_4",
           "coef_var_1_30", "coef_var_31_60",
           "gdd_wtr_0c.previous", "post_ice_warm_rate.previous", "date_over_16.7.previous", 
           "mean_surf_JulAugSep.previous","mean_surf_may.previous", "winter_dur_0_4.previous",
           "coef_var_1_30.previous", "coef_var_31_60.previous",
           "gdd_wtr_0c.2.previous", "post_ice_warm_rate.2.previous", "date_over_16.7.2.previous", 
           "mean_surf_JulAugSep.2.previous","mean_surf_may.2.previous", "winter_dur_0_4.2.previous",
           "coef_var_1_30.2.previous", "coef_var_31_60.2.previous"))

lattice::levelplot(cor)

mixed.rf.gn <- MixRF(Y = gn_temp_train$log_cpue , X = independent.gn, 
                  random = "(1|lake_id)", data = gn_temp_train, importance = T,
                  mtry = sqrt(ncol(independent.gn)), ntree = 500)
                  
#write_rds(mixed.rf.gn, "./models/adult.mixed.rf.rds")  
#
mixed.rf.gn <- readRDS("./models/adult.mixed.rf.rds")
#mixed.rf.gn$forest
#
varImpPlot(mixed.rf.gn$forest, type = 1)

random.gn <- as.data.frame(ranef(mixed.rf.gn$MixedModel)) %>%
  mutate(
    spot = "Spot",
    upper = condval + 2*condsd,
    lower = condval - 2*condsd,
    spot = ifelse(upper*-1 <0& lower*-1<0, "Bright Spot", spot ),
    spot = ifelse(upper*-1 >0& lower*-1>0, "Dark Spot", spot )) %>%
  rename(lake_id = grp) %>% select(!c("grpvar", "term"))
```


# Adult bright spots plots
```{r, eval = F}
location <- gn_temp_train %>% select(lake_id, lake_name.1, x, y) %>% 
  distinct(lake_id, .keep_all = T)

bright.spots.gn <- merge(x = random.gn, y = location, 
                         by = "lake_id", all.x = T) %>%
  mutate(size = ifelse(spot == "Spot", T, F))

mn_counties <- map_data("state", "Minnesota")

group.colors <- c(`Dark Spot` = "black", `Spot` = "blue", `Bright Spot` = "red")
group.size <- c(`small` = 1, `large` = 2)

ggplot() + 
  geom_polygon(data = mn_counties, aes(x = long, y = lat, group = group), 
               colour="black", fill="NA", lwd=1 )  + 
  # color changes state outline, fill changes state fill
  coord_fixed(1.3) +
  geom_point(data = bright.spots.gn, 
             aes(x=x, y=y, color = spot, alpha = size, size = size)) + 
  scale_color_manual(values = group.colors) +
  scale_alpha_discrete(range = c(1, .25), guide = 'none') +
  scale_size_discrete(range = c(2, 1), guide = 'none') +
  # fill changes lake colors
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  labs(title = "Gill net",
       #subtitle = "1356 Lakes, 564 Bright Spots, 50 Dark spots",
       x = "Longitude", y = "Latitude", color = "Spots") 
#
ggsave("./figures/gillnet_brightspots.jpeg")


temp <- gn_temp_train
# messing around
temp <- temp %>% group_by(lake_id) %>% summarize(mean = mean(cpue))

toplot <- merge(bright.spots.gn, temp, by = c("lake_id"))

toplot %>% 
  ggplot() +
  geom_point(aes(
    y = condval, x = log(mean), 
    color = spot, alpha = size, size = size)) +
  scale_color_manual(values = group.colors) +
  scale_alpha_discrete(range = c(1, .25), guide = 'none') +
  scale_size_discrete(range = c(2, 1), guide = 'none') +
  labs(title = "Gill net",
       subtitle = "",
       x = "Average gill net CPUE", y = "Deviation from expected", 
       color = "Spots")

```

# Adult plots and messing around
```{r, eval = F}
gn.importance <- importance(mixed.rf.gn$forest) 
gn.importance <- cbind("variable" = rownames(gn.importance), 
                       data.frame(gn.importance, row.names=NULL)) %>%
  rename(increase_MSE = X.IncMSE)

gn.importance %>% 
  ggplot(aes(x=reorder(variable, increase_MSE), y = increase_MSE)) +
  geom_point() + 
  geom_segment(aes(x = variable, xend = variable, y=0, yend = increase_MSE)) + 
  coord_flip() + xlab("Variable") + ylab("% increase in MSE") + ggtitle("Gill net importance") +
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 12))
  
#ggsave("./figures/gill.net.importance.jpeg")



gn_temp_train$predicted = predict.MixRF(mixed.rf.gn, 
                                        gn_temp_train, EstimateRE = T)

#fitted vs predicted
gn_temp_train %>%
  ggplot() + 
  geom_point(aes(x = log(cpue), y = predicted)) + 
  geom_abline (slope=1, linetype = "dashed", color="Red") + 
  ggtitle("Gill net predicted vs fitted") + 
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 12))
#ggsave("./figures/gill.net.predVSfit.jpeg")
```

# Gill net partial plots
```{r}

for (i in 1:ncol(independent.gn)){
  col_names = colnames(independent.gn)
  pd <- pdp::partial(mixed.rf.gn$forest, pred.var =col_names[i], train = independent.gn)
  
  pd %>% ggplot() +
    geom_point(aes(y = pd[,2], x = pd[,1])) +
    geom_line(aes(y = pd[,2], x = pd[,1])) +
    xlab(col_names[i]) + ylab("y-hat") +
    theme_classic()
  ggsave(filename = paste0("./figures/gill_net_partial_plots/", col_names[i], "_partial.jpeg"))
}

```


# remove big objects (Gill-net)
```{r}
rm(rf); rm(mixed.rf.gn)
```


# Size structure modeling
```{r}
independent <- size_temp_train %>% 
  select(!c("lake_id", "lake_name.1", "n_over500", "p_over500"))

cor <- Hmisc::rcorr(as.matrix(independent))$r

# high correlation (>.6)
# acres-shore miles, secchi has ~.4-.5 with all temp metrics
# practically all temp metrics
# good temp variables post-ice-warm rate, coef_var, date_over

independent <- size_temp_train %>% 
  select(c("year", "julian_day", "cpue_yep","cpue_np","cpue_cisco",
           "log_acres", "max_depth", "elevation", "BACI", 
           "three_year_secchi",
           "STOCKED", "STOCKED.previous", "STOCKED.2.previous", 
           "fry.pa", "fry.pa.previous", "fry.pa.2.previous",
           "gdd_wtr_0c", "post_ice_warm_rate", "date_over_16.7", 
           "mean_surf_JulAugSep","mean_surf_may", "winter_dur_0_4",
           "coef_var_1_30", "coef_var_31_60",
           "gdd_wtr_0c.previous", "post_ice_warm_rate.previous", "date_over_16.7.previous", 
           "mean_surf_JulAugSep.previous","mean_surf_may.previous", "winter_dur_0_4.previous",
           "coef_var_1_30.previous", "coef_var_31_60.previous",
           "gdd_wtr_0c.2.previous", "post_ice_warm_rate.2.previous", "date_over_16.7.2.previous", 
           "mean_surf_JulAugSep.2.previous","mean_surf_may.2.previous", "winter_dur_0_4.2.previous",
           "coef_var_1_30.2.previous", "coef_var_31_60.2.previous"))



# Fit Mixed Effects RF
mixed.rf.size <- MixRF(Y = size_temp_train$p_over500 , X = independent, 
                  random = "(1|lake_id)", data = size_temp_train, 
                  importance = T,
                  mtry = sqrt(ncol(independent)), ntree = 1000)
                  
#write_rds(mixed.rf.size, "./models/adult.mixed.size.rf.rds")  
#
mixed.rf.size <- readRDS("./models/adult.mixed.size.rf.rds")
#
mixed.rf.size$forest
#
varImpPlot(mixed.rf.size$forest)

random.size <- as.data.frame(ranef(mixed.rf.size$MixedModel)) %>%
  mutate(
    spot = "Spot",
    upper = condval + 1.96*condsd,
    lower = condval - 1.96*condsd,
    spot = ifelse(upper*-1 <0& lower*-1<0, "Bright Spot", spot ),
    spot = ifelse(upper*-1 >0& lower*-1>0, "Dark Spot", spot )) %>%
  rename(lake_id = grp) %>% select(!c("grpvar", "term"))
```

# Size bright spots plots
```{r, eval = F}
location <- size_temp_train %>% select(lake_id, lake_name.1, x, y) %>% 
  distinct(lake_id, .keep_all = T)

bright.spots.size <- merge(x = random.size, y = location, 
                         by = "lake_id", all.x = T) %>%
  mutate(size = ifelse(spot == "Spot", T, F))

mn_counties <- map_data("state", "Minnesota")

group.colors <- c(`Dark Spot` = "black", `Spot` = "blue", `Bright Spot` = "red")
group.size <- c(`small` = 1, `large` = 2)

ggplot() + 
  geom_polygon(data = mn_counties, aes(x = long, y = lat, group = group), 
               colour="black", fill="NA", lwd=1 )  + 
  # color changes state outline, fill changes state fill
  coord_fixed(1.3) +
  geom_point(data = bright.spots.size, aes(x=x, y=y, color = spot, alpha = size, size = size)) + 
  scale_color_manual(values = group.colors) +
  scale_alpha_discrete(range = c(1, .25), guide = 'none') +
  scale_size_discrete(range = c(2, 1), guide = 'none') +
  # fill changes lake colors
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  labs(title = "Gill net",
       subtitle = "1006 Lakes, 98 Bright Spots, 75 Dark spots",
       x = "Longitude", y = "Latitude", color = "Spots") 
#ggsave("./figures/size_brightspots.jpeg")
```

# Size plots and messing around
```{r, eval = F}
size.importance <- importance(mixed.rf.size$forest) 
size.importance <- cbind("variable" = rownames(size.importance), 
                       data.frame(size.importance, row.names=NULL)) %>%
  rename(increase_MSE = X.IncMSE)

size.importance %>% 
  ggplot(aes(x=reorder(variable, increase_MSE), y = increase_MSE)) +
  geom_point() + 
  geom_segment(aes(x = variable, xend = variable, y=0, yend = increase_MSE)) + 
  coord_flip() + xlab("Variable") + ylab("% increase in MSE") + 
  ggtitle("Proportion over 500mm importance") +
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 12))
#ggsave("./figures/size.importance.jpeg")

#partialPlot(mixed.rf.size, independent, BACI)
#partialPlot(mixed.rf.size, independent, cpue_yep)
#partialPlot(mixed.rf.size, independent, gdd_wtr_0c)

# Fitted vs actual
size_temp_train$predicted = predict.MixRF(mixed.rf.size, size_temp_train, EstimateRE = T)

size_temp_train %>%
  ggplot() + 
  geom_point(aes(x = p_over500, y = predicted)) + 
  geom_abline (slope=1, linetype = "dashed", color="Red") +
  ggtitle("YOY predicted vs fitted") + 
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 12))
#ggsave("./figures/size.predVSfit.jpeg")
```



# Electro fishing modeling
```{r}
independent <- yoy_surveys %>% 
  select(!c("lake_id", "cpue"))

cor <- Hmisc::rcorr(as.matrix(independent))$r


independent.yoy <- yoy_surveys %>% 
  select(c("year", "julian_day",
           "log_acres","fry.pa", "STOCKED", "BACI", "median_secchi", 
           "peak_temp", "coef_var_1_30", "coef_var_31_60",
           "gdd_wtr_0c", "post_ice_warm_rate", "date_over_16.7",
           "date_over_21", "ice_off_yday",
           "mean_surf_JulAugSep","mean_surf_may"))


mixed.rf.recruit <- MixRF(Y = yoy_surveys$log_cpue , X = independent.yoy, 
                  random = "(1|lake_id)", data = yoy_surveys,
                  importance = T,
                  mtry = sqrt(ncol(independent.yoy)), ntree = 1000)
#write_rds(mixed.rf.recruit, "./models/recruit.mixed.rf.rds")
#
mixed.rf.recruit <- readRDS("./models/recruit.mixed.rf.rds")
#mixed.rf.recruit$forest
#varImpPlot(mixed.rf.recruit$forest)

random.yoy <- as.data.frame(ranef(mixed.rf.recruit$MixedModel)) %>%
  mutate(
    spot = "Spot",
    upper = condval + 1.96*condsd,
    lower = condval - 1.96*condsd,
    spot = ifelse(upper*-1 <0& lower*-1 < 0, "Bright Spot", spot ),
    spot = ifelse(upper*-1 >0& lower*-1 > 0, "Dark Spot", spot )) %>%
  rename(lake_id = grp) %>% select(!c("grpvar", "term"))
```

# recruitment bright spots plots
```{r, eval = F}
location.yoy <- yoy_surveys %>% select(lake_id, x, y) %>% 
  distinct(lake_id, .keep_all = T)

bright.spots.yoy <- merge(x = random.yoy, y = location.yoy, 
                         by = "lake_id", all.x = T) %>%
  mutate(size = ifelse(spot == "Spot", T, F))

mn_counties <- map_data("state", "Minnesota")

group.colors <- c(`Dark Spot` = "black", `Spot` = "blue", `Bright Spot` = "red")
group.size <- c(`small` = 1, `large` = 2)

ggplot() + 
  geom_polygon(data = mn_counties, aes(x = long, y = lat, group = group), 
               colour="black", fill="NA", lwd=1 )  + 
  # color changes state outline, fill changes state fill
  coord_fixed(1.3) +
  geom_point(data = bright.spots.yoy, aes(x=x, y=y, color = spot, alpha = size, size = size)) + 
  scale_color_manual(values = group.colors) +
  scale_alpha_discrete(range = c(1, .25), guide = 'none') +
  scale_size_discrete(range = c(2, 1), guide = 'none') +
  # fill changes lake colors
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  labs(title = "YOY electrofishing",
       x = "Longitude", y = "Latitude", color = "Spots") 
#ggsave("./figures/recruitment_brightspots.jpeg")
```

# YOY plots and messing around
```{r, eval = F}
yoy.importance <- importance(mixed.rf.recruit$forest) 
yoy.importance <- cbind("variable" = rownames(yoy.importance), 
                       data.frame(yoy.importance, row.names=NULL)) %>%
  rename(increase_MSE = X.IncMSE)

yoy.importance %>% 
  ggplot(aes(x=reorder(variable, increase_MSE), y = increase_MSE)) +
  geom_point() + 
  geom_segment(aes(x = variable, xend = variable, y=0, yend = increase_MSE)) + 
  coord_flip() + xlab("Variable") + ylab("% increase in MSE") + 
  ggtitle("YOY electrofishing importance") +
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 12))
#ggsave("./figures/recruitment.importance.jpeg")

#partialPlot(mixed.rf.recruit, independent, BACI)
#partialPlot(mixed.rf.recruit$forest, independent.yoy, fry.pa)
#partialPlot(mixed.rf.recruit, independent, gdd_wtr_0c)


# Fitted vs actual
yoy_surveys$predicted = predict.MixRF(mixed.rf.recruit, yoy_surveys, EstimateRE = T)

#fitted vs predicted
yoy_surveys %>%
  ggplot() + 
  geom_point(aes(x = log(cpue), y = predicted)) + 
  geom_abline (slope=1, linetype = "dashed", color="Red") +
  ggtitle("YOY predicted vs fitted") + 
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 12))
#ggsave("./figures/recruitment.predVSfit.jpeg")

```


# remove big objects (recruits)
```{r}
rm(mixed.rf.recruit)
```



# Overlap (GN - YOY)
```{r}
overlap.yoy.gn <- merge(x = bright.spots.yoy,  y = bright.spots.gn, by = c("lake_id")) %>%
  mutate(sychrony = ifelse(spot.x == spot.y, "Yes", "No")) %>%
  rename(yoy_spot = spot.x, gn_spot = spot.y,
         x = x.x, y= y.x ) %>%
  dplyr::select(lake_id, yoy_spot, gn_spot, sychrony, y, x)


ggplot() + 
  geom_polygon(data = mn_counties, aes(x = long, y = lat, group = group), 
               colour="black", fill="NA", lwd=1 )  + 
  # color changes state outline, fill changes state fill
  coord_fixed(1.3) +
  geom_point(data = overlap.yoy.gn, aes(x=x, y=y, color = sychrony, alpha = sychrony, size = sychrony)) + 
  #scale_color_manual(values = group.colors) +
  scale_alpha_discrete(range = c(1, .25), guide = 'none') +
  scale_size_discrete(range = c(2, 1), guide = 'none') +
  # fill changes lake colors
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) + 
  labs(title = "Sychrony between gillnet and\n recruitment brightspots",
       subtitle = "155 Lakes, 109 Yes, 46 No",
       x = "Longitude", y = "Latitude", color = "Sychrony\nbetween spots") 
#ggsave("./figures/synchrony.gn.yoy.jpeg")
```

# Overlap (GN - Size)
```{r}
overlap.size.gn <- merge(x = bright.spots.size,  y = bright.spots.gn, by = c("lake_id")) %>%
  mutate(sychrony = ifelse(spot.x == spot.y, "Yes", "No")) %>%
  rename(size_spot = spot.x, gn_spot = spot.y,
         x = x.x, y= y.x ) %>%
  dplyr::select(lake_id, size_spot, gn_spot, sychrony, y, x)

ggplot() + 
  geom_polygon(data = mn_counties, aes(x = long, y = lat, group = group), 
               colour="black", fill="NA", lwd=1 )  + 
  # color changes state outline, fill changes state fill
  coord_fixed(1.3) +
  geom_point(data = overlap.size.gn, aes(x=x, y=y, color = sychrony, alpha = sychrony, size = sychrony)) + 
  #scale_color_manual(values = group.colors) +
  scale_alpha_discrete(range = c(1, .25), guide = 'none') +
  scale_size_discrete(range = c(2, 1), guide = 'none') +
  # fill changes lake colors
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) + 
  labs(title = "Sychrony between gillnet and\n size structure brightspots",
       subtitle = "1006 Lakes, 765 Yes, 241 No",
       x = "Longitude", y = "Latitude", color = "Sychrony\nbetween spots") 
#ggsave("./figures/synchrony.gn.size.jpeg")
```

# Overlap (YOY - Size)
```{r}
overlap.size.gn <- merge(x = bright.spots.size,  y = bright.spots.yoy, by = c("lake_id")) %>%
  mutate(sychrony = ifelse(spot.x == spot.y, "Yes", "No")) %>%
  rename(size_spot = spot.x, yoy_spot = spot.y,
         x = x.x, y= y.x ) %>%
  dplyr::select(lake_id, size_spot, yoy_spot, sychrony, y, x)

ggplot() + 
  geom_polygon(data = mn_counties, aes(x = long, y = lat, group = group), 
               colour="black", fill="NA", lwd=1 )  + 
  # color changes state outline, fill changes state fill
  coord_fixed(1.3) +
  geom_point(data = overlap.size.gn, aes(x=x, y=y, color = sychrony, alpha = sychrony, size = sychrony)) + 
  #scale_color_manual(values = group.colors) +
  scale_alpha_discrete(range = c(1, .25), guide = 'none') +
  scale_size_discrete(range = c(2, 1), guide = 'none') +
  # fill changes lake colors
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) + 
  labs(title = "Sychrony between YOY and\n size structure brightspots",
       subtitle = "105 Lakes, 87 Yes, 28 No",
       x = "Longitude", y = "Latitude", color = "Sychrony\nbetween spots") 
#ggsave("./figures/synchrony.yoy.size.jpeg")
```

# hot spots and bright spots
```{r}
hotspot.gn <- read.csv("./data/hotspots.csv") %>% 
  filter(method == "Gill-net") %>% select(lake_id,lat,long) %>%
  rename(x =long, y = lat) %>% mutate(spots = "Hot Spot")

bright.spots.gn <- bright.spots.gn %>% 
  select(lake_id, x, y, spot)

hot.bright.gn <- merge(hotspot.gn, bright.spots.gn, all = T)
```

